---
title: "r2dii-plot-0-2-0"
description: |
  New styling functions, new arguments to `plot_*()`, and bug fixes.
author:
  - name: Mauro Lepore
    url: https://github.com/maurolepore
base_url: https://2degreesinvesting.github.io/
date: 2021-10-18
output: 
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: true
categories:
  - r2dii
  - package
  - release
preview: preview.jpg
twitter:
  site: "@mauro_lepore"
  creator: "@mauro_lepore"
draft: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

r2dii-plot 0.2.0 is now on CRAN. This is a minor release with new features,
enhancements and bug fixes. This post focuses on the new features. For
everything else see the
[changelog](https://2degreesinvesting.github.io/r2dii.plot/news/).

You can install r2dii.plot from CRAN with:

```r
install.packages("r2dii.plot")
```

And you can use it with `library()`. Here we'll also use the ggplot2 package.

```{r}
library(ggplot2, warn.conflicts = FALSE)
library(r2dii.plot)
```

### New styling functions

A new family of functions help you apply 2DII colour palettes to any plot. 

Note this plot uses data and functions from the ggplot2 package exclusively.

```{r}
p <- ggplot(mpg) + geom_point(aes(displ, hwy, color = class))
p
```

You can now use r2dii.plot to apply 2DII's colour scale with:

```{r}
p + scale_colour_r2dii()
```

To discover other functions in the "2dii scale" family run
`?scale_colour_r2dii()` and navigate to the section "See also".

### New arguments to `plot_*()`

The `plot_*()` functions gained new arguments to replicate features of the
`qplot_*()` functions.

For example, compare these two plots. Note the shorter (5 years) time span and
nicer legend labels in the plot created with `qplot_*()`.

```{r}
data <- subset(sda, sector == "cement")

qplot_emission_intensity(data)

plot_emission_intensity(data)
```

You can now replicate those features of `qplot_*()` with the new arguments
`span_5yr` and `convert_label`:

```{r}
plot_emission_intensity(data, span_5yr = TRUE, convert_label = to_title)
```

## Acknowledgements

```{r Acknowledgements, echo=FALSE}
# bash
# git shortlog --summary tag-oldrel..tag-release
# For example: Who contributed to r2dii.match 0.0.2 (since 0.0.1)?
# git shortlog --summary 0.0.1..0.0.2

thank <- function(x) {
  x <- sort(x)
  paste0("[", x, "]", sprintf("(https://github.com/%s)", x), collapse = ", ")
}
contributors <- c(
  "AlexAxthelm",
  "maurolepore",
  "MonikaFu"
)
```

Thanks to [\@cjyetman](https://github.com/cjyetman) for [proposing the `scale_*()` functions](https://github.com/2DegreesInvesting/r2dii.plot/issues/28), to [\@georgeharris2deg](https://github.com/georgeharris2deg) for reporting bugs,
and to all contributors to this release (`r thank(contributors)`).
