---
title: "r2dii.data 0.2.0 is now on CRAN"
description: |
  Signifcant changes to `ald_demo`
author:
  - name: Jackson Hoffart
    url: https://github.com/jdhoffa
base_url: https://2degreesinvesting.github.io/
date: 2021-06-18
output: 
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: true
categories:
  - r2dii
  - package
  - release
preview: preview.jpg
twitter:
  site: "@jdhoffa"
  creator: "@jdhoffa"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

r2dii.data 0.2.0 is now on CRAN. This release fixes a number of bugs that you can learn about [here](https://2degreesinvesting.github.io/r2dii.data/news/index.html); this post shows enhancements and new features.

You can install it from CRAN with:

```r
install.packages("r2dii.data")
```

Then use them with:

```{r}
library(r2dii.data)
library(r2dii.match)
```

The most notable update in this release relates to changes in the demonstration dataset `ald_demo`. This data is meant to reflect a dataset that is provided to users by the organization [Asset Resolution](https://asset-resolution.com/).

### `ald_demo` gains a column, loses a column an has two renamed

`ald_demo` gains the new column `lei`, loses the column `is_ultimate_listed_owner` and has two columns renamed: 
* `id_company` becomes `company_id`
* `ald_emission_factor_unit` becomes `emission_factor_unit`

The new column `lei` contains fake legal entity identifiers for some companies in `ald_demo`: 

``` {r} 
head(unique(ald_demo$lei))
```

Changes to the other columns are simply cosmetic, and meant to better reflect the actual dataset that will be delivered to users. 

Since these fields are not currently used in the downstream packages `r2dii.match` and `r2dii.analysis`, it is not expected that this should negatively affect users in a significant way. 

### `ald_demo` gains a new sector `hdv`

`ald_demo` also gains datapoints relating to the newly added Heavy-Duty Vehicle sector (abbreviated to `hdv`). 

``` {r}
head(ald_demo[ald_demo$sector == "hdv",])
```