---
title: "PACTA Terminology Update: ALD becomes ABCD"
description: |
  Explaining the decision to change terminology from Asset-Level Data to 
  Asset-Based Company Data.
author:
  - name: Jackson Hoffart
    url: https://github.com/jdhoffa
date: 2022-03-02
output: 
  distill::distill_article:
  toc: true
  toc_depth: 3
  self_contained: true
categories:
  - methodology
preview: preview.png
twitter:
  site: "@jdhoffart"
  creator: "@jdhoffart"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

``` {r packages, echo=FALSE}
suppressPackageStartupMessages(library(dplyr))
library(r2dii.data)
library(gt)
```

The PACTA team has recently decided to update the terminology used to refer to a
crucial dataset. The dataset that was previously known as "Asset-Level Data"
(ALD) will henceforth be referred to as "Asset-Based Company Data" (ABCD). This 
note explains our reasoning. 

## Explanation

The PACTA methodology depends fundamentally on 3 input datasets:

- Portfolio data (e.g. portfolios, loanbooks, etc.)
- Climate change mitigation scenarios (e.g. International Energy Agency's - 
Sustainable Development Scenario )^[https://www.transitionmonitor.com/pacta-for-banks-2020/methodology-and-supporting-materials/]
- Data on the production of physical assets 

Historically, this last dataset was referred to as the "Asset-Level Data" (ALD). 
This name was a natural choice, since the underlying data describes the 
production of physical assets that are owned and operated by various companies 
globally.

An asset-level dataset might contain the following information:

``` {r, echo = FALSE}

ald <- tibble::tribble(
  ~asset_id,          ~asset_name, ~sector,     ~technology, ~year, ~production, ~production_unit, ~plant_location,
         1L,        "a wind farm", "power", "renewablescap", 2020L,        100L,             "MW",            "ES",
         1L,        "a wind farm", "power", "renewablescap", 2025L,        200L,             "MW",            "ES",
         2L, "a coal-fired plant", "power",       "coalcap", 2020L,        500L,             "MW",            "DE",
         2L, "a coal-fired plant", "power",       "coalcap", 2025L,        300L,             "MW",            "DE"
  )

ald %>% 
  gt() %>%
  tab_header(
    title = "Asset-Level Data"
  ) %>% 
  cols_label(
    asset_id = "Asset ID",
    asset_name = "Asset Name",
    sector = "Sector",
    technology = "Technology",
    year = "Year",
    production = "Capacity",
    production_unit = "Capacity Unit",
    plant_location = "Plant Location"
  )

```

In the table above, data is given out directly at the level of the physical 
asset.

However, while it is possible to run PACTA using data at the asset-level, the
back-end for both the PACTA for Investors online tool, and the free data-set
provided for the PACTA for Banks desktop tool have been pre-processed. The
actual data used has already been aggregated from the asset-level, to the
company-level (for all companies owning the assets):

``` {r, echo = FALSE}

company_data <- tibble::tribble(
  ~company_id,     ~name_company, ~sector,     ~technology, ~year, ~production, ~production_unit, ~plant_location,
  "3", "renewables inc.", "power", "renewablescap", 2013L, 11347.84804,             "MW",            "AS",
  "3", "renewables inc.", "power", "renewablescap", 2014L,  11744.7738,             "MW",            "AS",
  "3", "renewables inc.", "power", "renewablescap", 2015L, 12141.69955,             "MW",            "AS",
  "3", "renewables inc.", "power", "renewablescap", 2016L,  12538.6253,             "MW",            "AS",
  "3", "renewables inc.", "power", "renewablescap", 2017L, 12935.55106,             "MW",            "AS",
  "3", "renewables inc.", "power", "renewablescap", 2018L, 13332.47681,             "MW",            "AS"
)

company_data %>% 
  gt() %>%
  tab_header(
    title = "Asset-Based Company Data"
  ) %>% 
  cols_label(
    company_id = "Company ID",
    name_company = "Company Name",
    sector = "Sector",
    technology = "Technology",
    year = "Year",
    production = "Capacity",
    production_unit = "Capacity Unit",
    plant_location = "Plant Location"
  )

```

Notice that, while the data itself is based on information at the physical
asset-level, it has already been aggregated to the company level (by ownership
stake). In this sense, the data is not at the "Asset-Level" but rather the
"Company-Level". We therefor consider "Asset-Based Company Data" to be a more 
accurate name for the data provided. 
