---
title: "Onboarding to a defensive workflow"
description: |
  The minimum elements of a defensive workflow and links to learn more.
author:
  - name: Mauro Lepore
    url: https://github.com/maurolepore
base_url: https://2degreesinvesting.github.io/
date: 2021-10-26
output: 
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: true
categories:
  - community
  - onboarding
  - ds-incubator
preview: preview.jpg
twitter:
  site: "@mauro_lepore"
  creator: "@mauro_lepore"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

This article overviews the minimum elements of a defensive workflow and points
you to resources to learn more. Beyond onboarding to 2DII, this article should
be useful if you suspect you've drifted into habits that slow you down or make
your work more brittle (https://rstats.wtf/).

### [Setup RStudio to start from a blank slate every time you restart R](https://rstats.wtf/save-source.html)

Use these settings:

* Save workspace on quit: Never[^1].
* Load workspace on start: Off.

[^1]: Shift + Ctrl/Cmd + P
<img src=https://i.imgur.com/s1syyku.png width = 600>

Restart your R session often to confirm you can reproduce your results from a
blank slate.

### [Use RStudio projects](https://rstats.wtf/project-oriented-workflow.html#rstudio-projects):

RStudio makes it easy to create new projects[^2] and launch recent projects[^3].
To launch older projects it's easiest to use a dedicated app like Alfred[^4].

[^2]: File > New Project
[^3]: File > Recent Projects
[^4]: <img src=https://i.imgur.com/2TwTbpy.png width=500>

### [Practice safe paths](https://rstats.wtf/safe-paths.html)

Find your project's files with `here()`.

> Use it as a drop-in replacement for `file.path()`, it will always locate the
files relative to your project root.
-- https://here.r-lib.org/reference/here.html

```{r error=TRUE}
library(here)

# Good
(path <- here("_posts", "2021-10-26-defensive-workflow", "data", "greeting.txt"))
readLines(path)

# Bad
(path <- "/a/brittle path/that/only/i/have/data/greeting.txt")
readLines(path)
```

### [How to name files](https://rstats.wtf/how-to-name-files.html)

Name your files following these guidelines:

<img src=https://i.imgur.com/Lo7JFxc.png width=700>

https://speakerdeck.com/jennybc/how-to-name-files
