---
title: "r2dii.data 0.1.6, r2dii.match 0.0.7 and r2dii.analysis 0.1.3 are now on CRAN"
description: |
  Squashing bugs before the holidays.
author:
  - name: Jackson Hoffart
    url: https://github.com/jdhoffa
base_url: https://2degreesinvesting.github.io/
date: 2020-12-07
output: 
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: true
categories:
  - r2dii
  - package
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

r2dii.data 0.1.6, r2dii.match 0.0.7 and r2dii.analysis 0.1.3 are now on CRAN. These releases fix a number of bugs. All changes are listed in the changelog at the website of each package ([r2dii.data](https://2degreesinvesting.github.io/r2dii.data/news/index.html), [r2dii.match](https://2degreesinvesting.github.io/r2dii.match/news/index.html), [r2dii.analysis](https://2degreesinvesting.github.io/r2dii.analysis/news/index.html)); this post highlights the major bugs fixes and feature additions.

You can update or install these packages from CRAN using:

```r
install.packages("r2dii.data")
install.packages("r2dii.match")
install.packages("r2dii.analysis")
```

To use r2dii packages, you can load them into your active R session with `library()`. This example uses some other packages that you may also load now.

```{r}
library(dplyr, warn.conflicts = FALSE)
library(ggplot2)
library(r2dii.data)
library(r2dii.match)
library(r2dii.analysis)
```

### r2dii.data 0.1.6

Most changes to r2dii.data are with updates to the internal `region` and `classification` files. In particular, the `cnb_classification` dataset was added, to facilitate the sector classification of Nigerian banks: 

```{r}
cnb_classification %>% 
  filter(sector == "aviation") # show some illustrative results for aviation
```

The country code for Kosovo was also updated to the correct value of `xk`:

```{r}
filter(iso_codes, country == "kosovo")
```

### r2dii.match 0.0.7

The only user-facing change to r2dii.match was that `match_name()` gained the argument `...` to pass arguments to `stringdist::stringsim()`. Most users won't need this feature: 

```{r}
matched <- match_name(loanbook_demo, ald_demo)

matched %>% 
  filter(score != 1) %>% 
  select(name, name_ald, score)
```

However, advanced users can now specify many different parameters of the different methods available to `stringdist::stringsim`. For example, the `q` parameter of the `qgram` method can now be tweaked directly:

``` {r}
matched_qgram <- match_name(
  loanbook_demo,
  ald_demo, 
  method = "qgram",
  q = 1.5
)

matched_qgram %>% 
  filter(score != 1) %>% 
  select(name, name_ald, score)
```

For more details see `?stringdist::stringsim`.

### r2dii.analysis 0.1.3

There are a few major fixes in this release of r2dii.analysis, in particular to the functions `target_market_share` and `target_sda`. 

To explain those bugfixes, let's calculate some more demo data frames:

``` {r}
validated_matched <- matched %>% 
  prioritize()

# loan weighted portfolio-level results
portfolio_market_share_targets <- target_market_share(
  validated_matched,
  ald_demo,
  scenario_demo_2020,
  region_isos_demo
)

# unweighted company-level results
company_market_share_targets <- target_market_share(
  validated_matched,
  ald_demo,
  scenario_demo_2020,
  region_isos_demo,
  by_company = TRUE,
  weight_production = FALSE
)
```

#### target_market_share()

`target_market_share` outputs a data frame with the value `technology_share`. This value is expected to always sum to `1` when aggregated across appropriate groupings (e.g. by `sector`, `region`, `metric`). Before, this was true only for results at company level, but not portfolio level. Now, this is fixed and `technology_share` sums to `1` at both levels (at least to 10 significant digits; further digits may be off due to rounding errors):

```{r}
portfolio_sum <- portfolio_market_share_targets %>%
  group_by(sector, metric, region, year) %>%
  summarize(share_sum = sum(technology_share), .groups = "drop")

# true for all, to 10 significant digits
all(round(portfolio_sum$share_sum, 10) == 1)

company_sum <- company_market_share_targets %>%
  group_by(sector, metric, region, year, name_ald) %>%
  summarize(share_sum = sum(technology_share), .groups = "drop")

# true for all, to 10 significant digits
all(round(company_sum$share_sum, 10) == 1)
```

#### target_sda()

The function `target_sda` calculates a target pathway for `emission_factor` by computing a convergence target to the year 2050, as detailed in [this article](https://2degreesinvesting.github.io/r2dii.analysis/articles/target-sda.html). A bug in the function was causing this target to be calculated erroneously early, to the last year present in the input `ald` file. 

The convergence targets are now correctly calculated to the final year present in the input `scenario`:

``` {r}
matched <- tibble::tribble(
  ~id_loan, ~loan_size_outstanding, ~loan_size_outstanding_currency, ~loan_size_credit_limit, ~loan_size_credit_limit_currency, ~id_2dii, ~level, ~score, ~sector, ~name_ald, ~sector_ald,
  "L1", 1, "EUR", 2, "EUR", "UP1", "ultimate_parent", 1, "steel", "company a", "steel"
)

# an ALD file with values only between 2020 and 2030
ald <- tibble::tribble(
  ~name_company, ~sector, ~technology, ~year, ~production, ~emission_factor, ~plant_location, ~is_ultimate_owner,
  "company a", "steel", "steel", 2020, 1, 1.5, "DE", TRUE,
  "company a", "steel", "steel", 2025, 1, 1.5, "DE", TRUE,
  "company a", "steel", "steel", 2030, 1, 1.5, "DE", TRUE,
  "company b", "steel", "steel", 2020, 1, 2.5, "DE", TRUE,
  "company b", "steel", "steel", 2025, 1, 2.5, "DE", TRUE,
  "company b", "steel", "steel", 2030, 1, 2.5, "DE", TRUE
)

# a scenario file with targets at 2050
co2_scenario <- tibble::tribble(
  ~scenario_source, ~scenario, ~sector, ~region, ~year, ~emission_factor, ~emission_factor_unit,
  "etp_2017", "b2ds", "steel", "global", 2020, 2, "tonnes of CO2 per tonne of steel",
  "etp_2017", "b2ds", "steel", "global", 2025, 1.9, "tonnes of CO2 per tonne of steel",
  "etp_2017", "b2ds", "steel", "global", 2030, 1.8, "tonnes of CO2 per tonne of steel",
  "etp_2017", "b2ds", "steel", "global", 2050, 0.25, "tonnes of CO2 per tonne of steel",
)

# SDA portfolio-level results
sda_targets <- target_sda(
  matched,
  ald,
  co2_scenario
)

ggplot(
  data = sda_targets,
  mapping = aes(
    x = year,
    y = emission_factor_value,
    color = emission_factor_metric
  )
) +
  geom_line() +
  facet_wrap(~sector)
```

## Acknowledgements

```{r Acknowledgements, echo=FALSE}
# bash
# git shortlog --summary tag-oldrel..tag-release
# For example: Who contributed to r2dii.match 0.0.2 (since 0.0.1)?
# git shortlog --summary 0.0.1..0.0.2

thank <- function(x) {
  x <- sort(x)
  paste0("[", x, "]", sprintf("(https://github.com/%s)", x), collapse = ", ")
}
contributors <- c(
  "jdhoffa",
  "maurolepore"
)
```

While this release includes commits from only a few of us (`r thank(contributors)`), it is thanks to feedback from our colleagues and users.
